The Emperor Has No Clothes

### 2011-04-20 Winds/papers/LBG_Lya/lya_lbg.tex
jxp
git-svn-id: https://svn.ucolick.org/xsvn/trunk/radtrans@5912 ac11df5e-a788-c23b-a8e1-c758c323924d

\newcommand{\tmax}{$\tau_v^{\rm max}$}
\def\N#1{{N({\rm #1})}}
\newcommand{\mtmax}{\tau_v^{\rm max}}
\newcommand{\mtauv}{\tau_v}
\newcommand{\tauv}{$\tau_v$}
\newcommand{\mmc}{m_{\rm c}}
\def\mrvir{r_{\rm vir}}
\def\rvir{$r_{\rm vir}$}
\def \msun {M_\odot}
\def\wlya{$W^{\rm Ly\alpha}$}
\def\mwlya{W^{\rm Ly\alpha}}
\newcommand{\tpk}{$\tau_{\rm pk}$}
\newcommand{\mtpk}{\tau_{\rm pk}}
\newcommand{\mdv}{\delta v}
\newcommand{\ewabs}{$W_{\rm a}$}
\newcommand{\ewe}{$W_{\rm e}$}
\newcommand{\rtt}{$r_{\tau = 0.2}$}
\newcommand{\mrtt}{r_{\tau = 0.2}}
\newcommand{\bturb}{$b_{\rm turb}$}
\newcommand{\mbturb}{b_{\rm turb}}
\newcommand{\taud}{$\tau_{\rm dust}$}
\newcommand{\mtaud}{\tau_{\rm dust}}
\newcommand{\maconfig}{2p$^6$3s}
\newcommand{\mbconfig}{2p$^6$3p}
\newcommand{\aconfig}{a~$^6$D$^0$}
\newcommand{\zconfig}{z~$^6$D$^0$}
\newcommand{\mvr}{v_{\rm r}}
\newcommand{\naid}{\ion{Na}{1}~$\lambda\lambda 5891, 5897$}
\newcommand{\mgii}{\ion{Mg}{2}}
\newcommand{\mgiid}{\ion{Mg}{2}~$\lambda\lambda 2796, 2803$}
\newcommand{\mgiia}{\ion{Mg}{2}~$\lambda 2796$}
\newcommand{\mgiib}{\ion{Mg}{2}~$\lambda 2803$}
\newcommand{\feiia}{\ion{Fe}{2}~$\lambda 2586$}
\newcommand{\feiib}{\ion{Fe}{2}~$\lambda 2600$}
\newcommand{\feiic}{\ion{Fe}{2}$* \; \lambda 2612$}
\newcommand{\feiie}{\ion{Fe}{2}$* \; \lambda 2626$}
\newcommand{\feiid}{\ion{Fe}{2}~$\lambda\lambda 2586, 2600$}
\newcommand{\feiis}{\ion{Fe}{2}$*$}
\newcommand{\nmg}{$n_{\rm Mg^+}$}
\newcommand{\mnmg}{n_{\rm Mg^+}}
\newcommand{\nfe}{$n_{\rm Fe^+}$}
\newcommand{\mnfe}{n_{\rm Fe^+}}
\def\hub{h_{72}^{-1}}
\def\umfp{{\hub \, \rm Mpc}}
\def\mzq{z_q}
\def\zabs{$z_{\rm abs}$}
\def\mzabs{z_{\rm abs}}
\def\intl{\int\limits}
\def\cmma{\;\;\; ,}
\def\perd{\;\;\; .}
\def\ltk{\left [ \,}
\def\ltp{\left ( \,}
\def\ltb{\left \{ \,}
\def\rtk{\, \right  ] }
\def\rtp{\, \right  ) }
\def\rtb{\, \right \} }
\def\sci#1{{\; \times \; 10^{#1}}}
\def \rAA {\rm \AA}
\def \zem {$z_{\rm em}$}
\def \mzem {z_{\rm em}}
\def\smm{\sum\limits}
\def \cmm  {cm$^{-2}$}
\def \cmmm {cm$^{-3}$}
\def \kms  {km~s$^{-1}$}
\def \mkms  {\, {\rm km~s^{-1}}}
\def \lyaf {Ly$\alpha$ forest}
\def \Lya  {Ly$\alpha$}
\def \lya  {Ly$\alpha$}
\def \mlya  {Ly\alpha}
\def \Lyb  {Ly$\beta$}
\def \lyb  {Ly$\beta$}
\def \lyg  {Ly$\gamma$}
\def \ly5  {Ly-5}
\def \ly6  {Ly-6}
\def \ly7  {Ly-7}
\def \nhi  {$N_{\rm HI}$}
\def \mnhi  {N_{\rm HI}}
\def \lnhi {$\log N_{HI}$}
\def \mlnhi {\log N_{HI}}
\def \etal {\textit{et al.}}
\def \lyaf {Lyman--$\alpha$ forest}
\def \mnmin {\mnhi^{\rm min}}
\def \nmin {$\mnhi^{\rm min}$}
\def \O {${\mathcal O}(N,X)$}
\newcommand{\cm}[1]{\, {\rm cm^{#1}}}
\def \snrlim {SNR$_{lim}$}

@@

The LBG clump model proposed by S+10 envisions an isotropic outflow of
dense clumps, each optically thick to \lya\ and metal-line
transitions, but as an ensemble have a partial covering fraction to
the LBG and background sources.  We discuss in PKR11 our radiative
transfer treatment for this wind in the context of \mgiid.  In the
following, we adopt the same prescription but consider
\lya\footnote{Following standard practice, we treat \lya\ as a single
  transition instead of a double because the line separation is much less
  than the astrophysical velocities of interest here.}
and modify the parameters of the LBG clump model to those presented in
S+10.

In contrast to PKR11, we include two additional features in the
model.  First, we have explored two intrinsic emission models for the
galaxy near the \lya\ line: 
(i) a flat continuum, primarily generated by O and B stars within the
star-forming galaxy; and
(ii) a continuum plus \lya\ emission-line model parameteized by a
line-width $\sigma_{\rm Ly\alpha}$ and an equivalent width \wlya.
This line is presumed to be generated by \ion{H}{2} regions within the
galaxy.  We emphasize that neithre of these simple models are likely
to describe the complex emission of a star-forming galaxy at \lya.
Nevertheless, they help capture some of the salient features of the
LBG clump model.

Second, we have modeled attenuation of photons near \lya\ by the
intergalactic medium (IGM) surrounding LBGs on scales of $\sim 100\,
{\rm kpc} - 10$\,Mpc.  Our approach follows the method of
\cite{santos04}, using the \cite{barkana04} model for infall onto
galactic halos at $z \approx 3$.  We assume a galaxy with dark matter
halo mass $M_{\rm DM} \approx 10^{12} \msun$, virial radius $\mrvir 
= 80$\,kpc, circular velocity $v_{\rm c} = 220 \mkms$, and an
IGM with $T = 10^4$\,K, and a EUVB with ionization rate $\Gamma =
10^{-12.5} \, \rm s^{-1}$ \citep{fpl+08}.  With this parameterization,
we can calculate the density and velocity profiles of gas at $r >
\mrvir$ for the LBG (and assume the gas at $r<\mrvir$ is dominated by
the outflow) and apply the Sobolev approximation to calculate the
optical depth \tauv\ to \lya\ as a function of velocity.  We cuation that
this $\exp(-\mtauv)$ prescription does not capture the complexity of
gas on these scales.  Indeed, \cite{zheng10a} have argued that a more
proper treatment tends to yield higher opacity.
[INCLUDE THE BOOST?]
In either case, it is a necessary consideration for modeling the
observed \lya\ emission from these galaxies.

Figure~\ref{fig:1D} (top panels) presents the 
calculated 1D \lya\ spectrum for the
two intrinsic emission models of the galaxy.
The second panels show the predicted results when one includes
radiative transfer through the LBG clump model.  Similar to the \mgii\
results of PKR11, we observe significant absorption by the outflow for
$v < 0\mkms$.  The absorption actually peaks at $v < 0\mkms$ because
scattered emission has `filled-in' the absorption for $v \approx 0
\mkms$.  This emission peaks just redward of $v=0\mkms$, related to
photons emitted transverse to our sightline and scattered once to
Earth, and then extends to positive velocity, with a tail to $v
\approx +500\mkms$.  [Comment on Lya emission line model?]
The third pair of panels includes the effects of IGM scattering.  For
the DM halo and IGM model assumed, we find a peak optical depth
$\mtmax \approx 7$ at $v \approx v_{\rm c}$ that rapidly declines to
$\mtauv < 0.5$ at $v=0\mkms$.  This implies complete attenuation of
the \lya\ emission at $v \approx v_{\rm c}$ but only modest
attenuation at lower velocity.  For this parameterization, it is evident that
the effects of the IGM cannot be ignored.  If the IGM were just $5
\times$ more neutral than modeled here, the \lya\ line would be
completely ``extinguished'' at $v \gtrsim 0 \mkms$, as predicted for
the $z \sim 6$ universe \citep{zheng+10a}.  By the same token, if the
galaxy emits ionizing photons then \tmax\ may fall below unity and the
IGM would play a minor role.  Clearly, a full treatment of these
effects is warranted for the $z \sim 3$ universe.

In the bottom panels, we present the full model convolved with an
ideal spectrometer with FWHM~$\approx 250\mkms$ compared against the
observed, average \lya\ emission profile of LBGs from
\cite{shapley03}.  
At this spectral resolution,
typical of LBG spectroscopic observations, the weak emission that was
apparent for the continuum model is lost and one predicts a simple
absorption profile that extends to positive velocity because of IGM
attenuation.  This model is a very poor description of the data.  
For the line+continuum intrinsic model, in contrast, it is difficult
to detect the outflow and the IGM attenuation is completely lost at
this spectral resolution within the \lya\ emission profile.
Neither model predicts a significant emission line that is shifted by
several hundred \kms\ redward of systemic.  Indeed, reviewing the
results on \mgii\ from PKR11, such a shift is predicted for any
outflow model that we have considered.  The only scenario in PKR11
that showed a significant shift was when we included significant
opacity at $v \approx 0 \mkms$ at the source, i.e.\ an ISM model.
This has the the effect of altering the intrinsic emission [How?] 
and when combined with an outflow can yield a redshifted emission
profile.  [Do this with Lya!] 
[Comment that infall can produce a redshift \lya\ line too, i.e. Zheng
et al.]

[Previous authors have modeled the \lya\ profiles of LBGs and have
concluded the redshift emission is well decribed by an outflow
\citep[e.g.][]{verhomme}.  Similar to the
ISM+wind model described above, these outflow models only predict a
redshift if there is a very large opacity at $v \approx 0\mkms$.  In
the \cite{verhomme} models, this is accomplished by considering a
thin expanding shell with a total \ion{H}{1} column density $\mnhi > 10^{20}
\cm{-2}$.  In this case, the damping wings of the gas imply a large
opacity at $v \approx 0 \mkms$ provided the shell has a speed less
than $\approx 500 \mkms$.  This model, then, predicts a very strong
absorption profile at $v < 0 \mkms$, which is generally not observed
in the LBG spectra.  One of the few exceptions is the galaxy cb58.]

>>>>

\def\N#1{{N({\rm #1})}}
\newcommand{\tmax}{$\tau_v^{\rm IGM, max}$}
\newcommand{\mtmax}{\tau_v^{\rm IGM, max}}
\newcommand{\mtauv}{\tau_v}
\newcommand{\tauv}{$\tau_v$}
\newcommand{\tigm}{$\tau_v^{\rm IGM}$}
\newcommand{\mtigm}{\tau_v^{\rm IGM}}
\newcommand{\mmc}{m_{\rm c}}
\def\mrvir{r_{\rm vir}}
\def\rvir{$r_{\rm vir}$}
\def \msun {M_\odot}
\def\wlya{$W^{\rm Ly\alpha}$}
\def\mwlya{W^{\rm Ly\alpha}}
\newcommand{\tpk}{$\tau_{\rm pk}$}
\newcommand{\mtpk}{\tau_{\rm pk}}
\newcommand{\mdv}{\delta v}
\newcommand{\ewabs}{$W_{\rm a}$}
\newcommand{\ewe}{$W_{\rm e}$}
\newcommand{\rtt}{$r_{\tau = 0.2}$}
\newcommand{\mrtt}{r_{\tau = 0.2}}
\newcommand{\bturb}{$b_{\rm turb}$}
\newcommand{\mbturb}{b_{\rm turb}}
\newcommand{\taud}{$\tau_{\rm dust}$}
\newcommand{\mtaud}{\tau_{\rm dust}}
\newcommand{\maconfig}{2p$^6$3s}
\newcommand{\mbconfig}{2p$^6$3p}
\newcommand{\aconfig}{a~$^6$D$^0$}
\newcommand{\zconfig}{z~$^6$D$^0$}
\newcommand{\mvr}{v_{\rm r}}
\newcommand{\naid}{\ion{Na}{1}~$\lambda\lambda 5891, 5897$}
\newcommand{\mgii}{\ion{Mg}{2}}
\newcommand{\mgiid}{\ion{Mg}{2}~$\lambda\lambda 2796, 2803$}
\newcommand{\mgiia}{\ion{Mg}{2}~$\lambda 2796$}
\newcommand{\mgiib}{\ion{Mg}{2}~$\lambda 2803$}
\newcommand{\feiia}{\ion{Fe}{2}~$\lambda 2586$}
\newcommand{\feiib}{\ion{Fe}{2}~$\lambda 2600$}
\newcommand{\feiic}{\ion{Fe}{2}$* \; \lambda 2612$}
\newcommand{\feiie}{\ion{Fe}{2}$* \; \lambda 2626$}
\newcommand{\feiid}{\ion{Fe}{2}~$\lambda\lambda 2586, 2600$}
\newcommand{\feiis}{\ion{Fe}{2}$*$}
\newcommand{\nmg}{$n_{\rm Mg^+}$}
\newcommand{\mnmg}{n_{\rm Mg^+}}
\newcommand{\nfe}{$n_{\rm Fe^+}$}
\newcommand{\mnfe}{n_{\rm Fe^+}}
\def\hub{h_{72}^{-1}}
\def\umfp{{\hub \, \rm Mpc}}
\def\mzq{z_q}
\def\zabs{$z_{\rm abs}$}
\def\mzabs{z_{\rm abs}}
\def\intl{\int\limits}
\def\cmma{\;\;\; ,}
\def\perd{\;\;\; .}
\def\ltk{\left [ \,}
\def\ltp{\left ( \,}
\def\ltb{\left \{ \,}
\def\rtk{\, \right  ] }
\def\rtp{\, \right  ) }
\def\rtb{\, \right \} }
\def\sci#1{{\; \times \; 10^{#1}}}
\def \rAA {\rm \AA}
\def \zem {$z_{\rm em}$}
\def \mzem {z_{\rm em}}
\def\smm{\sum\limits}
\def \cmm  {cm$^{-2}$}
\def \cmmm {cm$^{-3}$}
\def \kms  {km~s$^{-1}$}
\def \mkms  {\, {\rm km~s^{-1}}}
\def \lyaf {Ly$\alpha$ forest}
\def \Lya  {Ly$\alpha$}
\def \lya  {Ly$\alpha$}
\def \mlya  {Ly\alpha}
\def \Lyb  {Ly$\beta$}
\def \lyb  {Ly$\beta$}
\def \lyg  {Ly$\gamma$}
\def \ly5  {Ly-5}
\def \ly6  {Ly-6}
\def \ly7  {Ly-7}
\def \nhi  {$N_{\rm HI}$}
\def \mnhi  {N_{\rm HI}}
\def \lnhi {$\log N_{HI}$}
\def \mlnhi {\log N_{HI}}
\def \etal {\textit{et al.}}
\def \lyaf {Lyman--$\alpha$ forest}
\def \mnmin {\mnhi^{\rm min}}
\def \nmin {$\mnhi^{\rm min}$}
\def \O {${\mathcal O}(N,X)$}
\newcommand{\cm}[1]{\, {\rm cm^{#1}}}
\def \snrlim {SNR$_{lim}$}

@@

The LBG clump model proposed by S+10 envisions an isotropic outflow of
dense clumps, each optically thick to \lya\ and metal-line
transitions, but which as an ensemble have a partial covering fraction to
the LBG and background sources.  We discuss in PKR11 our radiative
transfer treatment for this wind in the context of \mgiid.  In the
following, we adopt the same algorithms but focus on
\lya\footnote{Following standard practice, we treat \lya\ as a single
  transition instead of a double because the line separation is much less
  than the astrophysical velocities of interest here.}
and modify the parameters of the LBG clump model as appropriate.

In contrast to PKR11, we include two additional features in the
model.  First, we have explored two intrinsic emission models for the
galaxy near the \lya\ line: 
(i) a flat continuum, assumed to be generated by O and B stars within the
star-forming galaxy; and
(ii) a continuum plus \lya\ emission-line model parameterized by a
line-width $\sigma_{\rm Ly\alpha}$ and a \lya\ equivalent width \wlya.
The \lya\ emission is presumed to be generated by \ion{H}{2} regions within the
galaxy.  We emphasize that neither of these simple models are likely
to describe the complex emission of a star-forming galaxy at \lya.
Nevertheless, they help capture some of the salient features of the
LBG clump wind model.

Second, we have modeled the attenuation of photons near \lya\ by the
IGM surrounding LBGs on scales of $\sim 100\,
{\rm kpc} - 10$\,Mpc.  Our approach follows the method of
\cite{santos04}, using the \cite{barkana04} model for infall onto
galactic halos at $z \approx 3$.  We assume a galaxy with dark matter
halo mass $M_{\rm DM} \approx 10^{12} \msun$, virial radius $\mrvir 
= 80$\,kpc, circular velocity $v_{\rm c} = 220 \mkms$, and an
IGM with $T = 10^4$\,K.  THis amterial is bathed in a uniform
extragalactic UV
background (EUVB) radiation field with ionization rate $\Gamma =
10^{-12.5} \, \rm s^{-1}$ \citep{fpl+08}.  With this parameterization,
we can calculate the density and velocity profiles of gas at $r >
\mrvir$ for the LBG (and assume the gas at $r\le \mrvir$ is dominated by
the outflow) and apply the Sobolev approximation to calculate the
optical depth $\tigm$ to \lya\ as a function of velocity.
We caution that
this $\exp(-\mtigm)$ prescription does not capture the complexity of
gas on these scales.  Indeed, \cite{zheng10a} have argued that a more
proper treatment tends to yield even higher average opacity.
[INCLUDE THE BOOST? Yes, and justify to match $D_A$]
[CHECK against Djisktra]
Therefore, one may wish to consider our estimation of \tigm\ as a
lower limit.

Figure~\ref{fig:1D} (top panels) presents the 
calculated 1D \lya\ spectrum for the
two intrinsic emission models of the galaxy.
For the line-emission, we have adopted $\sigma_{\rm Ly\alpha} =
50\mkms$ and $\mwlya = 100$\AA, the latter of which is suggested for
actively star-forming galaxies \citep{lya_emit}.
The second panels show the predicted results when one includes
radiative transfer through the LBG clumpy wind model.  Similar to the \mgii\
results of PKR11, we observe significant absorption by the outflow for
$v < 0\mkms$.  
Although the opacity of the outflow peaks at $v \approx 0 \mkms$, the
minimum flux is actually observed to be closer to $v \approx -300
\mkms$ because
scattered emission `fills-in' the absorption at $v \lesssim 0
\mkms$.  Most of these scattered photons were
emitted transverse to our sightline and then scattered once to
Earth.  The tail of emission to positive velocities ($v
\approx +500\mkms$) in contrast, are associated with photons that
scattered off the far-side of this isotropic outflow.  
We emphasize that this emission is weak in comparison with the
flux at $v \approx 0 \mkms$, simply because of geometrical projection.

[Emphasize that in this clump model, photons don't diffuse in
frequency much]

The third pair of panels includes the effects of IGM scattering.  For
the DM halo and IGM model assumed, we find a peak optical depth
$\mtmax \approx 7$ at $v \approx v_{\rm c}$ that rapidly declines to
$\mtauv < 0.5$ at $v=0\mkms$.  This implies complete attenuation of
the \lya\ emission at $v \approx v_{\rm c}$ but only modest
attenuation at lower velocities.  Although the IGM effects for our
parameterization are relatively modest compared to the outflow, it is evident that
the IGM cannot be ignored at this redshift.  If the IGM were just $5
\times$ more neutral than modeled here, the \lya\ line would be
completely ``extinguished'' at $v \gtrsim 0 \mkms$, as predicted for
the $z \sim 6$ universe \citep{zheng+10a}.  By the same token, if the
galaxy emits ionizing photons then \tmax\ may fall below unity and the
IGM would play a minor role.  Clearly, a full treatment 
is warranted for the $z \sim 3$ universe.

In the bottom panels we present the full model, convolved with an
ideal spectrometer having FWHM~$\approx 250\mkms$, compared against the
average \lya\ emission profile observed for a sample of $z \sim 3$ 
LBGs \citep{shapley03}. 
[Match our resolution to Shapley??]
At this spectral resolution, which is
typical of LBG spectroscopic observations, the weak emission at $v
\approx 0 \mkms$ that was
apparent for the continuum model is lost and one predicts a simple
absorption profile that extends to positive velocity. 
This model is a very poor description of the data.  
For the line+continuum intrinsic model, in contrast, it is difficult
to detect the outflow and the IGM attenuation is completely lost at
this spectral resolution within the \lya\ emission profile.
Neither model predicts significant line-emission 
at several hundred \kms\ redward of systemic.  
Such emission is {\it not a natural prediction of a clumpy outflow
  model} where the photons tend to escape because of partial covering
not photon diffusion.  In fact, reviewing the
results on \mgii\ from PKR11, such shifts are rare in
outflow models that we have considered.  The only models in PKR11
with significant shifts to positive velocity was when there was large 
opacity at $v \approx 0 \mkms$ near the source, e.g.\ models that
included an ISM component. In essence, this gas modifies the intrinsic
line profile of \lya, yielding the standard `double hump' emission and
then the blue hump is strongly attenuated by the outflow.
[Test this with Lya!] 
It is also worth noting that a strict infall scenario on Mpc scales
also naturaly predicts a redshifted \lya\ line.  The IGM model of
\cite{zheng+10}, for example, predicts \lya\ emission will be detectd
at $v \approx +XX \mkms$ from star-forming galaxies.

In contrast to our assertions,
other authors have modeled the \lya\ profiles of LBGs and have
concluded the redshifted emission is a `clear signature' of an outflow
\citep[e.g.][]{verhomme}.  
[page 7]

Similar to the ISM+wind model described above, these outflow models only predict a
redshift if there is a very large opacity at $v \approx 0\mkms$.  In
the \cite{verhomme} models, this is accomplished by considering a
thin expanding shell with a total \ion{H}{1} column density $\mnhi > 10^{20}
\cm{-2}$.  In this case, the damping wings of the gas imply a large
opacity at $v \approx 0 \mkms$ provided the shell has a speed less
than $\approx 500 \mkms$.  This model, then, predicts a very strong
absorption profile at $v < 0 \mkms$, which is generally not observed
in the LBG spectra.  One of the few exceptions is the galaxy cb58.]
